### 2014-05-12 Images/markovdiagram.tex
added a lot of stuff

\begin{tikzpicture}[scale=.7, every node/.style={scale=0.7}]
    % -----------------------------------------------
    % Boundaries ------------------------------------
    % -----------------------------------------------
    \draw [dashed, line width=1mm, red, opacity=.3] (12,2) -- (12,-20);
    \draw [dashed, line width=1mm, red, opacity=.3] (-2,-12) -- (20,-12);
    %\tikzstyle{state}=[ellipse, draw, fill=blue!10, minimum width=2cm];
    %\tikzstyle{state}=[rectangle, draw, fill=blue!10, minimum width=2cm];
    \tikzstyle{state}=[minimum width=2cm, font=\boldmath];
    %\draw  (0,0) rectangle (4,3);
    %\draw [dashed, red, thick] (0,1) -- (4,1);
    %\draw [dashed, red, thick] (3,0) -- (3,3);
    % -----------------------------------------------
    % First row--------------------------------------
    % -----------------------------------------------
    \node (aa) [state] at (0,0) {$(0,0)$};
    \node (ab) [state] at ($(aa) + (3,0)$) {$(0,1)$};
    \node (ac) [state] at ($(ab) + (3,0)$) {$(0,2)$};
    \node (ellipsisa1) at ($(ac) + (3,0)$) {$\dots$};
    \node(ak) [state] at ($(ellipsisa1) + (3,0)$) {$(0,K_{\RG})$};
    \node (ellipsisa2) at ($(ak) + (3,0)$) {$\dots$};
    \node(az) [state] at ($(ellipsisa2) + (3,0)$) {$(0,c_{\RG})$};
    % Transitions -----------------------------------
    % On Row:
    \draw (aa) edge[out=45,in=135,->,thick] node [above] {\tiny$\lambda_{\RG}^{(a)}$} (ab);
    \draw (aa) edge[out=-45,in=-135,<-,thick] node [below] {\tiny$\mu_{\RG}$} (ab);

    \draw (ab) edge[out=45,in=135,->,thick] node [above] {\tiny$\lambda_{\RG}^{(a)}$} (ac);
    \draw (ab) edge[out=-45,in=-135,<-,thick] node [below] {\tiny$\mu_{\RG}$} (ac);

    \draw (ac) edge[out=45,in=135,->,thick] node [above] {\tiny$\lambda_{\RG}^{(a)}$} (ellipsisa1);
    \draw (ac) edge[out=-45,in=-135,<-,thick] node [below] {\tiny$\mu_{\RG}$} (ellipsisa1);

    \draw (ellipsisa1) edge[out=45,in=135,->,thick] node [above] {\tiny$\lambda_{\RG}^{(a)}$} (ak);
    \draw (ellipsisa1) edge[out=-45,in=-135,<-,thick] node [below] {\tiny$\mu_{\RG}$} (ak);

    \draw (ak) edge[out=45,in=135,->,thick] node [above] {\tiny$\lambda_{\RG}^{(b)}$} (ellipsisa2);
    \draw (ak) edge[out=-45,in=-135,<-,thick] node [below] {\tiny$\mu_{\RG}$} (ellipsisa2);

    \draw (ellipsisa2) edge[out=45,in=135,->,thick] node [above] {\tiny$\lambda_{\RG}^{(b)}$} (az);
    \draw (ellipsisa2) edge[out=-45,in=-135,<-,thick] node [below] {\tiny$\mu_{\RG}$} (az);
    % -----------------------------------------------
    % Second row--------------------------------------
    % -----------------------------------------------
    \node (ba) [state] at (0,-3) {$(1,0)$};
    \node (bb) [state] at ($(ba) + (3,0)$) {$(1,1)$};
    \node (bc) [state] at ($(bb) + (3,0)$) {$(1,2)$};
    \node (ellipsisb1) at ($(bc) + (3,0)$) {$\dots$};
    \node(bk) [state] at ($(ellipsisb1) + (3,0)$) {$(1,K_{\RG})$};
    \node (ellipsisb2) at ($(bk) + (3,0)$) {$\dots$};
    \node(bz) [state] at ($(ellipsisb2) + (3,0)$) {$(1,c_{\RG})$};
    % Transitions -----------------------------------
    % On Row:
    \draw (ba) edge[out=45,in=135,->,thick] node [above] {\tiny$\lambda_{\RG}^{(a)}$} (bb);
    \draw (ba) edge[out=-45,in=-135,<-,thick] node [below] {\tiny$\mu_{\RG}$} (bb);

    \draw (bb) edge[out=45,in=135,->,thick] node [above] {\tiny$\lambda_{\RG}^{(a)}$} (bc);
    \draw (bb) edge[out=-45,in=-135,<-,thick] node [below] {\tiny$\mu_{\RG}$} (bc);

    \draw (bc) edge[out=45,in=135,->,thick] node [above] {\tiny$\lambda_{\RG}^{(a)}$} (ellipsisb1);
    \draw (bc) edge[out=-45,in=-135,<-,thick] node [below] {\tiny$\mu_{\RG}$} (ellipsisb1);

    \draw (ellipsisb1) edge[out=45,in=135,->,thick] node [above] {\tiny$\lambda_{\RG}^{(a)}$} (bk);
    \draw (ellipsisb1) edge[out=-45,in=-135,<-,thick] node [below] {\tiny$\mu_{\RG}$} (bk);

    \draw (bk) edge[out=45,in=135,->,thick] node [above] {\tiny$\lambda_{\RG}^{(b)}$} (ellipsisb2);
    \draw (bk) edge[out=-45,in=-135,<-,thick] node [below] {\tiny$\mu_{\RG}$} (ellipsisb2);

    \draw (ellipsisb2) edge[out=45,in=135,->,thick] node [above] {\tiny$\lambda_{\RG}^{(b)}$} (bz);
    \draw (ellipsisb2) edge[out=-45,in=-135,<-,thick] node [below] {\tiny$\mu_{\RG}$} (bz);
    % With previous row:
    \draw (aa) edge[out=-125,in=125,->,thick] node [left] {\tiny$\lambda_{\NH}^{(a)}$} (ba);
    \draw (aa) edge[out=-55,in=55,<-,thick] node [right] {\tiny$\mu_{\NH}$} (ba);

    \draw (ab) edge[out=-125,in=125,->,thick] node [left] {\tiny$\lambda_{\NH}^{(a)}$} (bb);
    \draw (ab) edge[out=-55,in=55,<-,thick] node [right] {\tiny$\mu_{\NH}$} (bb);

    \draw (ac) edge[out=-125,in=125,->,thick] node [left] {\tiny$\lambda_{\NH}^{(a)}$} (bc);
    \draw (ac) edge[out=-55,in=55,<-,thick] node [right] {\tiny$\mu_{\NH}$} (bc);

    \draw (ellipsisa1) edge[out=-125,in=125,->,thick] node [left] {\tiny$\lambda_{\NH}^{(a)}$} (ellipsisb1);
    \draw (ellipsisa1) edge[out=-55,in=55,<-,thick] node [right] {\tiny$\mu_{\NH}$} (ellipsisb1);

    \draw (ak) edge[out=-125,in=125,->,thick] node [left] {\tiny$\lambda_{\NH}^{(b)}$} (bk);
    \draw (ak) edge[out=-55,in=55,<-,thick] node [right] {\tiny$\mu_{\NH}$} (bk);

    \draw (ellipsisa2) edge[out=-125,in=125,->,thick] node [left] {\tiny$\lambda_{\NH}^{(b)}$} (ellipsisb2);
    \draw (ellipsisa2) edge[out=-55,in=55,<-,thick] node [right] {\tiny$\mu_{\NH}$} (ellipsisb2);

    \draw (az) edge[out=-125,in=125,->,thick] node [left] {\tiny$\lambda_{\NH}^{(b)}$} (bz);
    \draw (az) edge[out=-55,in=55,<-,thick] node [right] {\tiny$\mu_{\NH}$} (bz);
    % -----------------------------------------------
    % Third row--------------------------------------
    % -----------------------------------------------
    \node (ca) [state] at (0,-6) {$(2,0)$};
    \node (cb) [state] at ($(ca) + (3,0)$) {$(2,1)$};
    \node (cc) [state] at ($(cb) + (3,0)$) {$(2,2)$};
    \node (ellipsisc1) at ($(cc) + (3,0)$) {$\dots$};
    \node(ck) [state] at ($(ellipsisc1) + (3,0)$) {$(2,K_{\RG})$};
    \node (ellipsisc2) at ($(ck) + (3,0)$) {$\dots$};
    \node(cz) [state] at ($(ellipsisc2) + (3,0)$) {$(2,c_{\RG})$};
    % Transitions -----------------------------------
    % On Row:
    \draw (ca) edge[out=45,in=135,->,thick] node [above] {\tiny$\lambda_{\RG}^{(a)}$} (cb);
    \draw (ca) edge[out=-45,in=-135,<-,thick] node [below] {\tiny$\mu_{\RG}$} (cb);

    \draw (cb) edge[out=45,in=135,->,thick] node [above] {\tiny$\lambda_{\RG}^{(a)}$} (cc);
    \draw (cb) edge[out=-45,in=-135,<-,thick] node [below] {\tiny$\mu_{\RG}$} (cc);

    \draw (cc) edge[out=45,in=135,->,thick] node [above] {\tiny$\lambda_{\RG}^{(a)}$} (ellipsisc1);
    \draw (cc) edge[out=-45,in=-135,<-,thick] node [below] {\tiny$\mu_{\RG}$} (ellipsisc1);

    \draw (ellipsisc1) edge[out=45,in=135,->,thick] node [above] {\tiny$\lambda_{\RG}^{(a)}$} (ck);
    \draw (ellipsisc1) edge[out=-45,in=-135,<-,thick] node [below] {\tiny$\mu_{\RG}$} (ck);

    \draw (ck) edge[out=45,in=135,->,thick] node [above] {\tiny$\lambda_{\RG}^{(b)}$} (ellipsisc2);
    \draw (ck) edge[out=-45,in=-135,<-,thick] node [below] {\tiny$\mu_{\RG}$} (ellipsisc2);

    \draw (ellipsisc2) edge[out=45,in=135,->,thick] node [above] {\tiny$\lambda_{\RG}^{(b)}$} (cz);
    \draw (ellipsisc2) edge[out=-45,in=-135,<-,thick] node [below] {\tiny$\mu_{\RG}$} (cz);
    % With previous row:
    \draw (ba) edge[out=-125,in=125,->,thick] node [left] {\tiny$\lambda_{\NH}^{(a)}$} (ca);
    \draw (ba) edge[out=-55,in=55,<-,thick] node [right] {\tiny$\mu_{\NH}$} (ca);

    \draw (bb) edge[out=-125,in=125,->,thick] node [left] {\tiny$\lambda_{\NH}^{(a)}$} (cb);
    \draw (bb) edge[out=-55,in=55,<-,thick] node [right] {\tiny$\mu_{\NH}$} (cb);

    \draw (bc) edge[out=-125,in=125,->,thick] node [left] {\tiny$\lambda_{\NH}^{(a)}$} (cc);
    \draw (bc) edge[out=-55,in=55,<-,thick] node [right] {\tiny$\mu_{\NH}$} (cc);

    \draw (ellipsisb1) edge[out=-125,in=125,->,thick] node [left] {\tiny$\lambda_{\NH}^{(a)}$} (ellipsisc1);
    \draw (ellipsisb1) edge[out=-55,in=55,<-,thick] node [right] {\tiny$\mu_{\NH}$} (ellipsisc1);

    \draw (bk) edge[out=-125,in=125,->,thick] node [left] {\tiny$\lambda_{\NH}^{(b)}$} (ck);
    \draw (bk) edge[out=-55,in=55,<-,thick] node [right] {\tiny$\mu_{\NH}$} (ck);

    \draw (ellipsisb2) edge[out=-125,in=125,->,thick] node [left] {\tiny$\lambda_{\NH}^{(b)}$} (ellipsisc2);
    \draw (ellipsisb2) edge[out=-55,in=55,<-,thick] node [right] {\tiny$\mu_{\NH}$} (ellipsisc2);

    \draw (bz) edge[out=-125,in=125,->,thick] node [left] {\tiny$\lambda_{\NH}^{(b)}$} (cz);
    \draw (bz) edge[out=-55,in=55,<-,thick] node [right] {\tiny$\mu_{\NH}$} (cz);
    % -----------------------------------------------
    % Fourth row--------------------------------------
    % -----------------------------------------------
    \node (da) at (0,-9) {$\vdots$};
    \node (db) at ($(da) + (3,0)$) {$\vdots$};
    \node (dc) at ($(db) + (3,0)$) {$\vdots$};
    \node (ellipsisd1) at ($(dc) + (3,0)$) {$\dots$};
    \node(dk) at ($(ellipsisd1) + (3,0)$) {$\vdots$};
    \node (ellipsisd2) at ($(dk) + (3,0)$) {$\dots$};
    \node(dz) at ($(ellipsisd2) + (3,0)$) {$\vdots$};
    % Transitions -----------------------------------
    % On Row:
    \draw (da) edge[out=45,in=135,->,thick] node [above] {\tiny$\lambda_{\RG}^{(a)}$} (db);
    \draw (da) edge[out=-45,in=-135,<-,thick] node [below] {\tiny$\mu_{\RG}$} (db);

    \draw (db) edge[out=45,in=135,->,thick] node [above] {\tiny$\lambda_{\RG}^{(a)}$} (dc);
    \draw (db) edge[out=-45,in=-135,<-,thick] node [below] {\tiny$\mu_{\RG}$} (dc);

    \draw (dc) edge[out=45,in=135,->,thick] node [above] {\tiny$\lambda_{\RG}^{(a)}$} (ellipsisd1);
    \draw (dc) edge[out=-45,in=-135,<-,thick] node [below] {\tiny$\mu_{\RG}$} (ellipsisd1);

    \draw (ellipsisd1) edge[out=45,in=135,->,thick] node [above] {\tiny$\lambda_{\RG}^{(a)}$} (dk);
    \draw (ellipsisd1) edge[out=-45,in=-135,<-,thick] node [below] {\tiny$\mu_{\RG}$} (dk);

    \draw (dk) edge[out=45,in=135,->,thick] node [above] {\tiny$\lambda_{\RG}^{(b)}$} (ellipsisd2);
    \draw (dk) edge[out=-45,in=-135,<-,thick] node [below] {\tiny$\mu_{\RG}$} (ellipsisd2);

    \draw (ellipsisd2) edge[out=45,in=135,->,thick] node [above] {\tiny$\lambda_{\RG}^{(b)}$} (dz);
    \draw (ellipsisd2) edge[out=-45,in=-135,<-,thick] node [below] {\tiny$\mu_{\RG}$} (dz);
    % With previous row:
    \draw (ca) edge[out=-125,in=125,->,thick] node [left] {\tiny$\lambda_{\NH}^{(a)}$} (da);
    \draw (ca) edge[out=-55,in=55,<-,thick] node [right] {\tiny$\mu_{\NH}$} (da);

    \draw (cb) edge[out=-125,in=125,->,thick] node [left] {\tiny$\lambda_{\NH}^{(a)}$} (db);
    \draw (cb) edge[out=-55,in=55,<-,thick] node [right] {\tiny$\mu_{\NH}$} (db);

    \draw (cc) edge[out=-125,in=125,->,thick] node [left] {\tiny$\lambda_{\NH}^{(a)}$} (dc);
    \draw (cc) edge[out=-55,in=55,<-,thick] node [right] {\tiny$\mu_{\NH}$} (dc);

    \draw (ellipsisc1) edge[out=-125,in=125,->,thick] node [left] {\tiny$\lambda_{\NH}^{(a)}$} (ellipsisd1);
    \draw (ellipsisc1) edge[out=-55,in=55,<-,thick] node [right] {\tiny$\mu_{\NH}$} (ellipsisd1);

    \draw (ck) edge[out=-125,in=125,->,thick] node [left] {\tiny$\lambda_{\NH}^{(b)}$} (dk);
    \draw (ck) edge[out=-55,in=55,<-,thick] node [right] {\tiny$\mu_{\NH}$} (dk);

    \draw (ellipsisc2) edge[out=-125,in=125,->,thick] node [left] {\tiny$\lambda_{\NH}^{(b)}$} (ellipsisd2);
    \draw (ellipsisc2) edge[out=-55,in=55,<-,thick] node [right] {\tiny$\mu_{\NH}$} (ellipsisd2);

    \draw (cz) edge[out=-125,in=125,->,thick] node [left] {\tiny$\lambda_{\NH}^{(b)}$} (dz);
    \draw (cz) edge[out=-55,in=55,<-,thick] node [right] {\tiny$\mu_{\NH}$} (dz);
    % -----------------------------------------------
    % Fifth row--------------------------------------
    % -----------------------------------------------
    \node (ea) [state] at (0,-12) {$(K_{\NH},0)$};
    \node (eb) [state] at ($(ea) + (3,0)$) {$(K_{\NH},1)$};
    \node (ec) [state] at ($(eb) + (3,0)$) {$(K_{\NH},2)$};
    \node (ellipsise1) at ($(ec) + (3,0)$) {$\dots$};
    \node(ek) [state] at ($(ellipsise1) + (3,0)$) {$(K_{\NH},K_{\RG})$};
    \node (ellipsise2) at ($(ek) + (3,0)$) {$\dots$};
    \node(ez) [state] at ($(ellipsise2) + (3,0)$) {$(K_{\NH},c_{\RG})$};
    % Transitions -----------------------------------
    % On Row:
    \draw (ea) edge[out=45,in=135,->,thick] node [above] {\tiny$\lambda_{\RG}^{(c)}$} (eb);
    \draw (ea) edge[out=-45,in=-135,<-,thick] node [below] {\tiny$\mu_{\RG}$} (eb);

    \draw (eb) edge[out=45,in=135,->,thick] node [above] {\tiny$\lambda_{\RG}^{(c)}$} (ec);
    \draw (eb) edge[out=-45,in=-135,<-,thick] node [below] {\tiny$\mu_{\RG}$} (ec);

    \draw (ec) edge[out=45,in=135,->,thick] node [above] {\tiny$\lambda_{\RG}^{(c)}$} (ellipsise1);
    \draw (ec) edge[out=-45,in=-135,<-,thick] node [below] {\tiny$\mu_{\RG}$} (ellipsise1);

    \draw (ellipsise1) edge[out=45,in=135,->,thick] node [above] {\tiny$\lambda_{\RG}^{(c)}$} (ek);
    \draw (ellipsise1) edge[out=-45,in=-135,<-,thick] node [below] {\tiny$\mu_{\RG}$} (ek);

    \draw (ek) edge[out=45,in=135,->,thick] node [above] {\tiny$\lambda_{\RG}^{(d)}$} (ellipsise2);
    \draw (ek) edge[out=-45,in=-135,<-,thick] node [below] {\tiny$\mu_{\RG}$} (ellipsise2);

    \draw (ellipsise2) edge[out=45,in=135,->,thick] node [above] {\tiny$\lambda_{\RG}^{(d)}$} (ez);
    \draw (ellipsise2) edge[out=-45,in=-135,<-,thick] node [below] {\tiny$\mu_{\RG}$} (ez);
    % With previous row:
    \draw (da) edge[out=-125,in=125,->,thick] node [left] {\tiny$\lambda_{\NH}^{(a)}$} (ea);
    \draw (da) edge[out=-55,in=55,<-,thick] node [right] {\tiny$\mu_{\NH}$} (ea);

    \draw (db) edge[out=-125,in=125,->,thick] node [left] {\tiny$\lambda_{\NH}^{(a)}$} (eb);
    \draw (db) edge[out=-55,in=55,<-,thick] node [right] {\tiny$\mu_{\NH}$} (eb);

    \draw (dc) edge[out=-125,in=125,->,thick] node [left] {\tiny$\lambda_{\NH}^{(a)}$} (ec);
    \draw (dc) edge[out=-55,in=55,<-,thick] node [right] {\tiny$\mu_{\NH}$} (ec);

    \draw (ellipsisd1) edge[out=-125,in=125,->,thick] node [left] {\tiny$\lambda_{\NH}^{(a)}$} (ellipsise1);
    \draw (ellipsisd1) edge[out=-55,in=55,<-,thick] node [right] {\tiny$\mu_{\NH}$} (ellipsise1);

    \draw (dk) edge[out=-125,in=125,->,thick] node [left] {\tiny$\lambda_{\NH}^{(b)}$} (ek);
    \draw (dk) edge[out=-55,in=55,<-,thick] node [right] {\tiny$\mu_{\NH}$} (ek);

    \draw (ellipsisd2) edge[out=-125,in=125,->,thick] node [left] {\tiny$\lambda_{\NH}^{(b)}$} (ellipsise2);
    \draw (ellipsisd2) edge[out=-55,in=55,<-,thick] node [right] {\tiny$\mu_{\NH}$} (ellipsise2);

    \draw (dz) edge[out=-125,in=125,->,thick] node [left] {\tiny$\lambda_{\NH}^{(b)}$} (ez);
    \draw (dz) edge[out=-55,in=55,<-,thick] node [right] {\tiny$\mu_{\NH}$} (ez);
    % Sixth row--------------------------------------
    \node (fa) at (0,-15) {$\vdots$};
    \node (fb) at ($(fa) + (3,0)$) {$\vdots$};
    \node (fc) at ($(fb) + (3,0)$) {$\vdots$};
    \node (ellipsisf1) at ($(fc) + (3,0)$) {$\dots$};
    \node(fk) at ($(ellipsisf1) + (3,0)$) {$\vdots$};
    \node (ellipsisf2) at ($(fk) + (3,0)$) {$\dots$};
    \node(fz) at ($(ellipsisf2) + (3,0)$) {$\vdots$};
    % Transitions -----------------------------------
    % On Row:
    \draw (fa) edge[out=45,in=135,->,thick] node [above] {\tiny$\lambda_{\RG}^{(c)}$} (fb);
    \draw (fa) edge[out=-45,in=-135,<-,thick] node [below] {\tiny$\mu_{\RG}$} (fb);

    \draw (fb) edge[out=45,in=135,->,thick] node [above] {\tiny$\lambda_{\RG}^{(c)}$} (fc);
    \draw (fb) edge[out=-45,in=-135,<-,thick] node [below] {\tiny$\mu_{\RG}$} (fc);

    \draw (fc) edge[out=45,in=135,->,thick] node [above] {\tiny$\lambda_{\RG}^{(c)}$} (ellipsisf1);
    \draw (fc) edge[out=-45,in=-135,<-,thick] node [below] {\tiny$\mu_{\RG}$} (ellipsisf1);

    \draw (ellipsisf1) edge[out=45,in=135,->,thick] node [above] {\tiny$\lambda_{\RG}^{(c)}$} (fk);
    \draw (ellipsisf1) edge[out=-45,in=-135,<-,thick] node [below] {\tiny$\mu_{\RG}$} (fk);

    \draw (fk) edge[out=45,in=135,->,thick] node [above] {\tiny$\lambda_{\RG}^{(d)}$} (ellipsisf2);
    \draw (fk) edge[out=-45,in=-135,<-,thick] node [below] {\tiny$\mu_{\RG}$} (ellipsisf2);

    \draw (ellipsisf2) edge[out=45,in=135,->,thick] node [above] {\tiny$\lambda_{\RG}^{(d)}$} (fz);
    \draw (ellipsisf2) edge[out=-45,in=-135,<-,thick] node [below] {\tiny$\mu_{\RG}$} (fz);
    % With previous row:
    \draw (ea) edge[out=-125,in=125,->,thick] node [left] {\tiny$\lambda_{\NH}^{(c)}$} (fa);
    \draw (ea) edge[out=-55,in=55,<-,thick] node [right] {\tiny$\mu_{\NH}$} (fa);

    \draw (eb) edge[out=-125,in=125,->,thick] node [left] {\tiny$\lambda_{\NH}^{(c)}$} (fb);
    \draw (eb) edge[out=-55,in=55,<-,thick] node [right] {\tiny$\mu_{\NH}$} (fb);

    \draw (ec) edge[out=-125,in=125,->,thick] node [left] {\tiny$\lambda_{\NH}^{(c)}$} (fc);
    \draw (ec) edge[out=-55,in=55,<-,thick] node [right] {\tiny$\mu_{\NH}$} (fc);

    \draw (ellipsise1) edge[out=-125,in=125,->,thick] node [left] {\tiny$\lambda_{\NH}^{(c)}$} (ellipsisf1);
    \draw (ellipsise1) edge[out=-55,in=55,<-,thick] node [right] {\tiny$\mu_{\NH}$} (ellipsisf1);

    \draw (ek) edge[out=-125,in=125,->,thick] node [left] {\tiny$\lambda_{\NH}^{(d)}$} (fk);
    \draw (ek) edge[out=-55,in=55,<-,thick] node [right] {\tiny$\mu_{\NH}$} (fk);

    \draw (ellipsise2) edge[out=-125,in=125,->,thick] node [left] {\tiny$\lambda_{\NH}^{(d)}$} (ellipsisf2);
    \draw (ellipsise2) edge[out=-55,in=55,<-,thick] node [right] {\tiny$\mu_{\NH}$} (ellipsisf2);

    \draw (ez) edge[out=-125,in=125,->,thick] node [left] {\tiny$\lambda_{\NH}^{(d)}$} (fz);
    \draw (ez) edge[out=-55,in=55,<-,thick] node [right] {\tiny$\mu_{\NH}$} (fz);
    % Final row--------------------------------------
    \node (za) [state] at (0,-18) {$(c_{\NH},0)$};
    \node (zb) [state] at ($(za) + (3,0)$) {$(c_{\NH},1)$};
    \node (zc) [state] at ($(zb) + (3,0)$) {$(c_{\NH},2)$};
    \node (ellipsisz1) at ($(zc) + (3,0)$) {$\dots$};
    \node(zk) [state] at ($(ellipsisz1) + (3,0)$) {$(c_{\NH},K_{\RG})$};
    \node (ellipsisz2) at ($(zk) + (3,0)$) {$\dots$};
    \node(zz) [state] at ($(ellipsisz2) + (3,0)$) {$(c_{\NH},c_{\RG})$};
    % Transitions -----------------------------------
    % On Row:
    \draw (za) edge[out=45,in=135,->,thick] node [above] {\tiny$\lambda_{\RG}^{(d)}$} (zb);
    \draw (za) edge[out=-45,in=-135,<-,thick] node [below] {\tiny$\mu_{\RG}$} (zb);

    \draw (zb) edge[out=45,in=135,->,thick] node [above] {\tiny$\lambda_{\RG}^{(d)}$} (zc);
    \draw (zb) edge[out=-45,in=-135,<-,thick] node [below] {\tiny$\mu_{\RG}$} (zc);

    \draw (zc) edge[out=45,in=135,->,thick] node [above] {\tiny$\lambda_{\RG}^{(d)}$} (ellipsisz1);
    \draw (zc) edge[out=-45,in=-135,<-,thick] node [below] {\tiny$\mu_{\RG}$} (ellipsisz1);

    \draw (ellipsisz1) edge[out=45,in=135,->,thick] node [above] {\tiny$\lambda_{\RG}^{(d)}$} (zk);
    \draw (ellipsisz1) edge[out=-45,in=-135,<-,thick] node [below] {\tiny$\mu_{\RG}$} (zk);

    \draw (zk) edge[out=45,in=135,->,thick] node [above] {\tiny$\lambda_{\RG}^{(d)}$} (ellipsisz2);
    \draw (zk) edge[out=-45,in=-135,<-,thick] node [below] {\tiny$\mu_{\RG}$} (ellipsisz2);

    \draw (ellipsisz2) edge[out=45,in=135,->,thick] node [above] {\tiny$\lambda_{\RG}^{(d)}$} (zz);
    \draw (ellipsisz2) edge[out=-45,in=-135,<-,thick] node [below] {\tiny$\mu_{\RG}$} (zz);
    % With previous row:
    \draw (fa) edge[out=-125,in=125,->,thick] node [left] {\tiny$\lambda_{\NH}^{(c)}$} (za);
    \draw (fa) edge[out=-55,in=55,<-,thick] node [right] {\tiny$\mu_{\NH}$} (za);

    \draw (fb) edge[out=-125,in=125,->,thick] node [left] {\tiny$\lambda_{\NH}^{(c)}$} (zb);
    \draw (fb) edge[out=-55,in=55,<-,thick] node [right] {\tiny$\mu_{\NH}$} (zb);

    \draw (fc) edge[out=-125,in=125,->,thick] node [left] {\tiny$\lambda_{\NH}^{(c)}$} (zc);
    \draw (fc) edge[out=-55,in=55,<-,thick] node [right] {\tiny$\mu_{\NH}$} (zc);

    \draw (ellipsisf1) edge[out=-125,in=125,->,thick] node [left] {\tiny$\lambda_{\NH}^{(c)}$} (ellipsisz1);
    \draw (ellipsisf1) edge[out=-55,in=55,<-,thick] node [right] {\tiny$\mu_{\NH}$} (ellipsisz1);

    \draw (fk) edge[out=-125,in=125,->,thick] node [left] {\tiny$\lambda_{\NH}^{(d)}$} (zk);
    \draw (fk) edge[out=-55,in=55,<-,thick] node [right] {\tiny$\mu_{\NH}$} (zk);

    \draw (ellipsisf2) edge[out=-125,in=125,->,thick] node [left] {\tiny$\lambda_{\NH}^{(d)}$} (ellipsisz2);
    \draw (ellipsisf2) edge[out=-55,in=55,<-,thick] node [right] {\tiny$\mu_{\NH}$} (ellipsisz2);

    \draw (fz) edge[out=-125,in=125,->,thick] node [left] {\tiny$\lambda_{\NH}^{(d)}$} (zz);
    \draw (fz) edge[out=-55,in=55,<-,thick] node [right] {\tiny$\mu_{\NH}$} (zz);

\end{tikzpicture}